A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-11-03 kobakian-cook.tex
Colours of centroid location

% !TeX root = RJwrapper.tex
\title{A Hexagon Tile Map Algorithm for Displaying Spatial Data}
\author{by Stephanie Kobakian and Dianne Cook}

@@

@@

A grid is created to ensure tessellation between the hexagons that
represent the geographic units on a hexagon tile map. The grid of
possible hexagon locations is made using the \texttt{create\_grid()}
function. It uses the centroids, the hexagon size and the buffer
distance.

@@

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \texttt{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess}a) shows the
hexagon grid after every second latitude row on the grid is shifted
right, by half of the hexagon size.

@@

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \texttt{create\_grid} function will call the
\texttt{create\_buffer} function. It finds the grid points needed to
best capture the set of centroids on a hexagon tile map. It also helps
reduce the amount of time taken, as it decreases the amount of points
considered for each centroid allocation.

@@

The hexagon grid points are kept only if they fall within the rolling
average of the minimum and maximum centroid values after accounting for
the buffer distance, for each row and column of the grid. Figure 5b
displays remaining hexagon grid points after applying the buffer step.
The sparsely populated South-West region of National Park has very few
points available compared to the South-East region near the city of
Hobart.

@@

The distance is calculated between each centroid in the set, and each of
the focal points provided. The distance and angle from the closest focal
point to polygon centroid is used in the allocation process as the order
for allocation is determined by the distance between the polygon
centroid and it's closest focal point. For this example, this distance
is only calculated for the capital city of Hobart.

@@

Allocation of all centroids takes place using the set of polygon
centroids and the hexagon map grid. The points are arranged from the
centroid closest to the focal point(s), to the furthest. Centroid
allocation begins with the closest centroid to a focal point. This will
preserve spatial relationships with the focal point, as the inner city
areas are allocated first, they will be placed closest to the capital,
and the areas that are further will then be accommodated. The set of
possible hexagon grid points reduces by one after each centroid is
allocated, then only the grid points that have not yet been selected are
considered.

The possible hexagon locations to consider for a centroid are determined
by the \texttt{hex\_filter}. This is the maximum amount of hexagons
between the centroid and the furthest considered hexagon. It is used to
subset possible grid points to only those surrounding the polygon
centroid within an appropriate range. A smaller distance will increase
speed, but can decrease accuracy when width of the angle increases.

The following example considers the first of the Statistical Areas at
Level 2. Within the algorithm, these steps are repeated for each
polygon.

@@

Keeping only the available hexagon points prevents multiple geographic
units from being allocated to the same hexagon.

@@

A box of possible hexagon locations around the centroid allows only the
closest points that are not yet assigned to be considered. The corners
of the box may not appear square if the buffer step has already removed
unnecessary points from over the ocean.

@@

The \texttt{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:filterprocess} . This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded, only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points.\\
By default the angle filter to hexagon grid points that fall within the
bounds of the angle from the focal point to the geographic centroid,
plus and minus 30 degrees. This will increase if no points can be found
within the \texttt{hex\_filter} distance. The default angle of 30 was
chosen to allow the algorithm to choose hexagons that best maintained
the spatial relationship between the focal point and geographic
centroid.

A complete hexagon tile map of Tasmania is created by applying the
algorithm steps to each centroid. The hexagon tile map is used to
visualise the Australian Cancer Atlas data \citep{TACA}. Two views of
the same data are produced by filling according to the thyroid Cancer
Standardised Incidence Rates (SIRs) downloaded from the Australian
Cancer Atlas site. This small example in Figure \ref{fig:filterprocess}
shows the group of blue areas in the Hobart CBD more prominently in the
hexagon tile map (b). The small red areas visible in the choropleth map
(a) along the north coast are much larger in the hexagon tile maps. The
hexagon tile map shows less yellow, this no longer overwhelms the map
space with the information regarding the rural areas.

When the algorithm is applied to the complete set of all Australian
Statistical Areas at Level 2 it results in the hexagon tile map display
shown in Figure \ref{fig:hexmap}. This display highlights the density of
Australian capital cities, as it draws attention to the many communities
in Sydney, Melbourne and Hobart. This display also highlights the
disparity in the burden of thyroid cancer for females in the communities
of these cities. There are several collections of red hexagons in Sydney
that represent the communities with much higher rates of diagnosis than
the Australian average. Brisbane also experiences higher than average
rates of diagnosis, but has more orange than red, which represents much
higher rates. The females in the cities of Adelaide and Perth show much
lower rates of diagnosis.

@@

>>>>

% !TeX root = RJwrapper.tex
\title{A Hexagon Tile Map Algorithm for Displaying Spatial Data}
\author{by Stephanie Kobakian, Dianne Cook, and Earl Duncan}

@@

When the algorithm is applied to the complete set of all Australian
Statistical Areas at Level 2 it results in the hexagon tile map display
shown in Figure \ref{fig:hexmap}. This display highlights the density of
Australian capital cities, as it draws attention to the many communities
in Sydney, Melbourne and Hobart. This display also highlights the
disparity in the burden of thyroid cancer for females in the communities
of these cities. There are several collections of red hexagons in Sydney
that represent the communities with much higher rates of diagnosis than
the Australian average. Brisbane also experiences higher than average
rates of diagnosis, but has more orange than red. The females in the
cities of Adelaide and Perth show much lower rates of diagnosis.

@@

A grid is created to allow tessellation of the hexagons that represent
the geographic units. For a hexagon tile map, the grid of possible
hexagon locations is made using the \texttt{create\_grid()} function. It
uses the centroids, the hexagon size and the buffer distance. Figure
\ref{fig:filterprocess}a) shows the initial locations of grid points
created for Tasmania.

@@

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \texttt{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess}a shows the
hexagon grid after every second latitude row on the grid is shifted
right, by half of the hexagon size.

@@

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \texttt{create\_grid} function will call the
\texttt{create\_buffer} function. It finds the grid points needed to
best capture the set of centroids on a hexagon tile map. Reducing the
buffer can help to decrease the amount of time taken, as there will be
less points in the water to consider for each centroid allocation.

@@

The hexagon grid points are kept only if they fall within the rolling
average of the minimum and maximum centroid values after accounting for
the buffer distance, for each row and column of the grid. Figure
\ref{fig:filterprocess}b displays remaining hexagon grid points after
applying the buffer filter. The sparsely populated South-West region of
National Park has fewer points available in the water compared to the
South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of
the focal points provided. The order for allocation is determined by the
distance between the polygon centroid and it's closest focal point. For
this example, this distance is only calculated for the capital city of
Hobart, represented in Figure \ref{fig:filterprocess}c and d as a white
cross.

@@

Allocation of all centroids requires the set of polygon centroids and
the hexagon map grid. The polygon centroids are ordered from the
centroid closest to the focal point(s), to the furthest. This will
preserve spatial relationships with the focal point, as the inner city
areas are allocated first and placed closest to the capital, the areas
that are further will then be accommodated. The following example
considers the first of the Statistical Areas at Level 2. Within the
algorithm, these steps are repeated for each polygon.

@@

The set of possible hexagon grid points reduces by one after each
centroid is allocated, then only the grid points that have not yet been
selected are considered. Keeping only the available hexagon points
prevents multiple geographic units from being allocated to the same
hexagon. This is demonstrated in Figure \ref{fig:filterprocess}c and d
by the black hexagons that represent the seven closest polygons to the
capital city of Hobart. As the allocation process begins for the eighth
closest centroid there are seven unavailable hexagon locations.

@@

A selection of possible hexagon locations around the centroid allows
only the closest points that are not yet assigned to be considered. The
number of possible hexagon locations to consider for a centroid is
determined by the \texttt{hex\_filter}. This is the maximum amount of
hexagons between the centroid and the furthest considered hexagon. It is
used to subset possible grid points to only those surrounding the
polygon centroid within an appropriate range. A smaller distance will
increase speed, but can decrease accuracy when width of the angle
increases.

@@

The \texttt{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:filterprocess}c.~This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded, only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points. By default the angle filter for the hexagon grid points
create plus and minus 30 degrees bounds of the angle from the focal
point to the geographic centroid. This will increase if no points can be
found within the \texttt{hex\_filter} distance. The default angle of 30
was chosen to allow the algorithm to find hexagons that best maintained
the spatial relationship between the focal point and geographic
centroid.

@@

\address{%
Earl Duncan\\
Queensland University of Technology\\%
School of Mathematical Sciences\\
%
%
%
\\\href{mailto:earl.duncan@qut.edu.au}{\nolinkurl{earl.duncan@qut.edu.au}}
}
